---
title: "Homework 3 (STAT 5860)"
author: "Eric Pettengill"
date: "Due by 11:59 pm, Feb. 24, 2018"
output:
  pdf_document: default
  word_document: default
---

## Instructions:

1. Download the `Homework3.Rmd` file from the course Elearning.

2. Open `Homework3.Rmd` in RStudio.

3. Replace the *"Your Name Here"* text in the `author` with your own name.

4. Write your answer to each problem by editing `Homework3.Rmd`.

5. After you finish all the problems, click `Knit to PDF` to create a pdf file. Upload your pdf file to Homework 1 Dropbox in the course Elearning.

### Set the seed number
```{r}
set.seed(224)
```

### Problem 1. The Rayleigh distribution is used to model lifetime subject to rapid aging, because the hazard rate is linearly increasing. The Rayleigh density is
$$
f(x) = \frac{x}{\sigma^2}e^{-x^2/(2\sigma^2)}, \quad x \geq 0, \sigma > 0.
$$

###### (a) Use the Metropolis-Hastings algorithm to generate a random sample of size 10000 from the Rayleigh $(\sigma = 2)$ distribution. Use the proposal distribution $\mbox{Gamma}(X_{t}, 1)$ (shape parameter $X_{t}$ and rate parameter 1). (Hint: Use \texttt{rgamma()} and \texttt{dgamma()})
```{r}
n <- 10000
x <- rep(0,n)

rayliegh <- function(x) {
  (x/4)*exp(-x^2/8)
}

q <- function(x) {
  rgamma(1, shape = x, rate = 1)
}

x[1] <- 5

for (i in 1:n) {
  xp <- q(x[i])
  alpha <- min(1, (rayliegh(xp)*dgamma(x[i], 
                                       shape = xp, 
                                       rate = 1))/(rayliegh(x[i])*dgamma(xp, 
                                                                         shape = x[i], 
                                                                         rate = 1)))
  if (runif(1) < alpha){
		x[i+1] <- xp
	}
	else{
		x[i+1] <- x[i]
	}	
}

```

###### (b) Plot the density histogram of the sample and add the theoretical density curve $f(x)$ to the density histogoram.
```{r, fig.width = 5, fig.height = 4}
hist(x, prob = TRUE, xlab = "", ylab = "", breaks = 50)
lines(seq(0,10,.01), rayliegh(seq(0,10,.01)))
```

###### (c) Find the Monte Carlo estimate of $\mathrm{E}_{f}(X)$ and draw a plot to show the convergence of the MCMC approximation.
```{r, fig.width = 4, fig.height = 4}
mean(x)


n <- 10000
q <- rgamma(n, shape = 5, rate = 1)

plot(cumsum(q*rayliegh(q)/dgamma(q, shape = 5, rate = 1))/(1:n),
     ylab = "",
     type = "l")
```

### Problem 2. The Cauchy distribution has density function
$$
f(x) = \frac{1}{\theta\pi\left[1 + \left(\frac{x-\eta}{\theta}\right)^2\right]}, \quad -\infty < x < \infty, \theta > 0.
$$

###### (a) Use the Metropolis algorithm to generate a random sample of size 10000 from the Cauchy distribuiton $(\theta = 2, \eta = 0)$. Here $\theta$ is scale parameter and $\eta$ is location parameter.
```{r}
n <- 10000
x <- rep(0,n)

cauch <- function(x) {
  1/(2*pi*(1+(x^2/4)))
}

q <- function(x) {
  rnorm(1, mean = x, sd = 4)
}

x[1] <- 0

for (i in 1:n) {
  xp <- q(x[i])
  alpha <- min(1, cauch(xp)/(cauch(x[i])))
  if (runif(1) < alpha){
		x[i+1] <- xp
	}
	else{
		x[i+1] <- x[i]
	}
}

```

###### (b) Compute the sample percentiles and compare with the Cauchy distribution $(\theta = 2, \eta = 0)$ percentiles. (Hint: Use \texttt{qcauchy()})
```{r, fig.width = 4, fig.height = 4}
p <- seq(0.1,0.9, 0.01)
Qhat <- quantile(x,p)
Q <- qcauchy(p, scale = 2, location = 0)

plot(Q,Qhat)
```


### Problem 3. Let $X$ = the number of heads obtained from flipping a fair coin four times. Here, we have that there are 4 trials, ``success'' is defined as getting a head face-up with probability of 0.5 for each trial, and the outcome of one trial doesn't affect another. 

###### (a) Use the Metropolis algorithm to simulate $X$, 50000 times. 
```{r}
n <- 50000
x <- rep(0,n)
mass <- dbinom(0:4, 4, .5)

S <- c(1:5)
Q <- matrix(rep(.2, 25), nrow = 5, ncol = 5)

x[1] <- 2

for(i in 1:n){
	xp <- sample(S, 1, prob = Q[x[i], ])
	alpha <- min(1, mass[xp]/mass[x[i]])
	if (runif(1) < alpha){
		x[i+1] <- xp
	}
	else{
		x[i+1] <- x[i]
	}
}

table(x)/(n+1)
```

###### (b) Find the Monte Carlo estimate of $\mathrm{E}(X)$.

$$E(X) = np = 4(.5) = 2$$

```{r}
n <- 50000
u <- runif(n)

z <- rbinom(u, 4, .5)

mean(z)

```



###### (c) Find the Monte Carlo estimate of $\mbox{Var}(X)$. (Hint: $\mbox{Var}(X) = \mathrm{E}(X^2) - (\mathrm{E}(X))^2$)

$$Var(X) = np(1-p) = 4(.5)(.5) = 1$$

```{r}
n <- 50000
u <- runif(n)

z <- rbinom(u, 4, .5)

mean((z - mean(z))^2)
```

