---
title: "Homework 2 (STAT 5860)"
author: "Eric Pettengill"
date: "Due by 11:59 pm, Feb. 12, 2018"
output:
  pdf_document: default
  word_document: default
---

## Instructions:

1. Download the `Homework2.Rmd` file from the course Elearning.

2. Open `Homework2.Rmd` in RStudio.

3. Replace the *"Your Name Here"* text in the `author` with your own name.

4. Write your answer to each problem by editing `Homework2.Rmd`.

5. After you finish all the problems, click `Knit to PDF` to create a pdf file. Upload your pdf file to Homework 1 Dropbox in the course Elearning.

### Set the seed number
```{r}
set.seed(212)
```

### Problem 1. The Pareto distribution has been used in economics as a model for a density function with a slowly decaying tail. The Pareto($a$,$b$) distribution has pdf
$$
f(x) = \frac{ab^{a}}{x^{a+1}}, \quad x \geq b
$$

### where $a > 0$ is shape paraemter and $b > 0$ is scale parameter. 

###### (a) Use the inverse transform method to generate a random sample of size 1000 from the Pareto(3,2) distribution.
```{r}
n <- 1000
u <- runif(n)
x <- 2/(1-u)^(1/3)
```

###### (b) Plot the density histogram of the sample.
```{r, fig.width = 5, fig.height = 4}
hist(x, probability = TRUE, xlim = c(2,10), breaks = 500)

```

###### (c) Add the theoretical density curve $f(x)$ to the density histogoram.
```{r, fig.width = 5, fig.height = 4}
hist(x, probability = TRUE, xlim = c(2,10), breaks = 500)

y <- seq(2, 30, .1)
lines(y, 24*y^(-4))

```


### Problem 2. A discrete random variable $X$ has probability mass functon (pmf)
  $x$  |   0   |   1   |   2   |   3   | 
------ | ----- | ----- | ----- | ----- |
$p(x)$ | 0.064 | 0.288 | 0.432 | 0.216 |

###### (a) Use the inverse transform method to generate a random sample of size 1000 from the distribution of $X$. Write your R code with \texttt{if} \texttt{else} statements. (Don't copy \texttt{while} loop code from Lecture 5 R code to solve this problem.)
```{r}
n <- 1000
x <- rep(0, n)
u <- runif(n)

for(i in 1:n){
  if ((u[i] <= .064) & (u[i] > 0)){
    x[i] <- 0 
  } else if ((u[i] <= .352) & (u[i] > .064)){
    x[i] <- 1
  } else if ((u[i] <= .784) & (u[i] > .352)){
    x[i] <- 2
  } else if ((u[i] <= 1) & (u[i] > .784)){
    x[i] <- 3
  }
}


```

###### (b) Calculate relative frequency for each $x$ and compare with the theoretical probabilities.
```{r, fig.width = 4, fig.height = 4}

for (i in 0:3) {
  print(sum(x == i)/length(x))
  }

```



### Problem 3. The normal distribution is one of the most important distribution in statistics because the nature provides numerous examples of populations of measurements that, at least approximately, follow a normal distribution. 

###### (a) Use the rejection sampling method to generate a random sample of size 1000 from the standard normal distribution. Use a standard Cauchy distribution as a proposal distribution $g(y)$. You can generate a random sample from the standard Cauchy distribution using \texttt{rcauchy()}. Moreover, you can calculate normal density and Cauchy density using \texttt{dnorm()} and \texttt{dcauchy()} respectively. Choose $c$ such that $f(y) \leq cg(y)$ and report your choice of $c$.  

Let $c = 2$.

```{r}
n <- 1000
k <- 0  ##ACCEPTED
j <- 0  ##ITERATIONS
x <- rep(0,n)

while (k < n){
  u <- runif(1)
  j <- j + 1
  cauch <- rcauchy(1)
  if ((dnorm(cauch)/(2*dcauchy(cauch))) > u){
    k <- k + 1
    x[k] <- cauch
  }
}


```

###### (b) Compute the sample percentiles and compare with the standard normal distribution percentiles.
```{r, fig.width = 4, fig.height = 4}
p <- seq(0.1,0.9, 0.01)
Qhat <- quantile(x,p)
Q <- qnorm(p)

plot(Q,Qhat)
```


### Problem 4. The random variable $X$ has pdf 
$$
f(x) = \frac{1}{2}\frac{1}{\sqrt{2\pi}}e^{-\frac{(x-1)^{2}}{2}} + \frac{1}{2}\frac{1}{\sqrt{2\pi}}e^{-\frac{(x-3)^{2}}{2}}, \quad -\infty < x < \infty
$$

###### (a) Use the importance sampling method to estimate $\mathrm{E}_{f}(X)$.
```{r}
n <- 10000
x <- rnorm(n, mean = 2, sd = 4)
y <- x*.5*(dnorm(x, mean = 1, sd = 1) + dnorm(x, mean = 3, sd = 1))/dnorm(x, mean = 2, sd = 4)
mean(y)
```

###### (b) Draw a plot to show the convergence of the importance sampling approximation.
```{r, fig.width = 4, fig.height = 4}
plot(cumsum((x*.5*(dnorm(x, mean = 1, sd = 1) + 
                  dnorm(x, mean = 3, sd = 1))/dnorm(x, mean = 2, sd = 4)))/(1:n), 
     ylab ="", 
     type = "l", 
     ylim = c(0.5,3))
```
